On the minimum leaf conductance g~0~

### 2017-03-29 manuscript.Rmd
Manuscript organized properly

---
title: "On g~0~"
author: ''
date: ''
output:
  html_document:
    theme: flatly
  word_document:
    reference_docx: manuscriptstyle.docx
csl: global-change-biology.csl
bibliography: references.bib
---

@@

```{r echo=FALSE}
includefigs <- F
library(png)
library(grid)

@@

# new outline

1. Introduction (short): questions about g0.
2. Need for inclusion of g0 in models
3. Comparison of values for various possible definitions
4. Estimation of g~0~ by fitting 
5. Nighttime conductance (v. short)
6. Cuticular conductance (short)
7. Minimum conductance (longer)
8. Recommendations and future work

# Abstract

# Introduction

@@

# Methods

## Data

## Model

# Do we need g~0~ in simulations of water and carbon fluxes?

@@

 **Cite Figure 1**

We need a non-zero minimum conductance when simulating water fluxes in very dry soil [@blackman2016; @anderegg2012]
Also @gleason2014

# g~0~ is poorly estimated from regression

@@

Barnard and Bauerle said so too but gave no details.
Even in very good datasets, g0 and g1 are strongly correlated (Figure 2).
Across datasets, g0 increases when the model fits poorly (Figure 3), and the width of the confidence interval increases when less variation in a ball-berry coefficient.
Often g0 is even negative [@heroult2013; @leuning1990], so it is often set to some arbitrary value

@@

g~0~ fitted is higher than minimum g~s~ from datasets. If we restric the fittings to those were g~0~ had a relatively narrow width of the confidence interval (< 0.03 mol m-2 s-1), then the average was `r with(subset(lin2015coef, g0_uci - g0_lci < 0.03), mean(g0))` compared to `r mean(lin2015coef$g0)` for all.

**Cite Figure 2 and 3**

# Definitions of minimum conductance

- gcut (cuticle without stomata)
- gmin (minimum conductance in situ, stomata closed)
- g0 (minimum conductance in the Ball-Berry-type model)
- gdark (nighttime conductance)

gdark > gmin > gcut
g0?

**Cite Figure 4**

@waldencoleman2013 Very interesting! gdark is different from gmin. Whole-plant WUE correlated with gdark (but not gmin). Lots of references on gdark and gmin and their variation. Closed system to measure gdark+gmin; gmin on detached leaves.

@muchow1989 correlation stomatal density and gmin suggesting leaky stomata

## Cuticular permeance

gseal on average not different from gcut - but for individual studies they frequently differ a lot indicating variable role of stomatal contribution (santrucek)

## Night-time conductance

Stomata simply do not close all the way, might stay open regardless of photosynthesis, or are broken.

## Minimum conductance from leaf-dry down curves

Hygen 1951 early work showing MLD method

true cuticular
MLD method (Figure 6) gives higher values but this is probably more useful for models (since these are whole leaves)

santrucek Direct measurements showing much higher 'cuticular' conductance of the leaf side with stomata, though stomata were closed. In agreement with @kerstiens1996 who showed 10x higher gmin for stomatal side vs. non-stomatal cuticle.

@burghardt2003 gmin was the same as cuticular permeance (Table 3) (Gmin is lowest gs in desiccating leaves, which includes cuticle and closed stomata). This demonstrates that stomata were able to completely close, and not lose significant amounts of water (with the exception of Hedera helix).

conductance: cuticle + closed stomata + cracks, broken stomata

- Kerstiens
@boyer1997 shows that water but not CO~2~ diffuses across non-stomatal cuticles.

- gmin as a plastic trait. Temperature response generally unresolved [@kerstiens2006]
- @riederer2001 correlation between temperature and gcut (non-stomatal cuticles)
- @schuster2016 correlation temperature, gcut AND gmin for Nerium oleander (page 59). T range much wider than Blackman though... But Schuster gmin not different from gcut at most temperatures, which is a bit strange too.
page 90 few gmin values from drying curves that we can add.

- @burghardt2003 gmin review
- @schreiber1996 cuticle review

[@anfodillo2002] gmin high altitude. Four-fold increase gmin from February (winter) to August in Picea abies, probably due to incomplete stomatal closure not changes in cuticle properties. Cite common finding that gmin increases with altitude (top left page 485).

[@herrick1991] gmin increased with altitude in Picea rubens.

[@jordan2007] large increase gmin in the herb Agastachys odorata with leaf age

@@

** cite figures 5,6 and 7**

# Future directions

## Figures

@@

@@

>>>>

---
title: "On the minimum leaf conductance g~0~"
author: ''
date: ''
output:
  html_document:
    theme: flatly
    number_sections: true
  word_document:
    reference_docx: manuscriptstyle.docx
csl: global-change-biology.csl
bibliography: references.bib
---

@@

```{r echo=FALSE}
includefigs <- T
library(png)
library(grid)

@@

# Introduction: questions about g0.

@@

# Need for inclusion of g0 in models

@@

We need a non-zero minimum conductance when simulating water fluxes in very dry soil [@blackman2016; @anderegg2012]
Also @gleason2014, @sinclair2000

# Comparison of values for various definitions

**Table 1.** Definitions of minimum conductance

| Variable   |  Definition |
| ---------- | ----------- |
| g~cut~     | Conductance of an isolated non-stomatal cuticle         |
| g~min~     | Minimum conductance measured from a leaf dry-down curve |
| g~dark~    | Night-time conductance                                  |
| g~0~       | Intercept in the Ball-Berry-type stomatal conductance model, fitted. |

@waldencoleman2013 gdark is much higher than gmin. Whole-plant WUE correlated with gdark (but not gmin). Lots of references on gdark and gmin and their variation. Closed system to measure gdark+gmin; gmin on detached leaves.

# Estimation of g~0~ by fitting

@@

[@barnard2013] mention difficulty of fitting but give no details.

Even in very good datasets, g0 and g1 are strongly correlated (Figure 2).
Across datasets, g0 increases when the model fits poorly (Figure 3), and the width of the confidence interval increases when less variation in a ball-berry coefficient.

Often g0 is even negative [@heroult2013; @leuning1990], so it is often set to some arbitrary value

@@

g~0~ fitted is higher than minimum g~s~ from datasets. If we restrict the fittings to those were g~0~ had a relatively narrow width of the confidence interval (< 0.03 mol m-2 s-1), then the average was `r with(subset(lin2015coef, g0_uci - g0_lci < 0.03), mean(g0))` compared to `r mean(lin2015coef$g0)` for all.

Relatively few reports on the plasticity of g~0~, perhaps because it is so poorly estimated. Exceptions:

@@

@@

# Nighttime conductance (v. short)

- gdark is variable during the night
- gdark is an active process not just a fixed parameter
- gs very early morning is quite high because in many cases gs increases before the sun comes up
- reviewed extensively elsewhere

#  Cuticular conductance (short)

- true cuticular conductance is that of isolated non-stomatal peels, and is very low compared to other definitions
- what is it, and is it variable with anything? yes, temperature, humidity.

Details of cuticular composition, modes of transport, and composition has been reviewed 
[@kerstiens1996; @kerstiens2006; @schreiber1996; @schreiber2001]

- @riederer2001 correlation between temperature and gcut (non-stomatal cuticles)

# Minimum conductance (longer)

[@hygen1951] early work showing MLD method

- similar to minimum gs observed for the Lin dataset
- plastic trait

One reason gmin is higher than gcut is the role of stomata, even when presumably shut or as shut as they will be.

gseal on average not different from gcut - but for individual studies they frequently differ a lot indicating variable role of stomatal contribution [@santrucek2004; @brodribb2014]. @muchow1989 correlation stomatal density and gmin suggesting leaky stomata. [@santrucek2004] Direct measurements showing much higher conductance of the leaf side with stomata, though stomata were closed.

but not always, in @burghardt2003 gmin was the same as cuticular permeance (Gmin is lowest gs in desiccating leaves, which includes cuticle and closed stomata). This demonstrates that stomata were able to completely close, and not lose significant amounts of water (with the exception of Hedera helix).

## Minimum conductance acclimates to growing conditions

@anfodillo2002 gmin high altitude. Four-fold increase gmin from February (winter) to August in Picea abies, probably due to incomplete stomatal closure not changes in cuticle properties. Cite common finding that gmin increases with altitude (top left page 485).

@herrick1991 gmin increased with altitude in Picea rubens.

Hypothesis that conifers have low gmin may have to be revisited in the light of @brodribb2014 (check what they actually say!)

@jordan2007 large increase gmin in the herb Agastachys odorata with leaf age

- @schuster2016 correlation temperature, gcut AND gmin for Nerium oleander (page 59). T range much wider than Blackman though... But Schuster gmin not different from gcut at most temperatures, which is a bit strange too.
page 90 few gmin values from drying curves that we can add.

@@

# Recommendations and future work

g0 at the moment covers everything, but we need separate ones (night, high VPD/T, drought, low PAR during day)
g0 is poorly estimated from regression - we need to figure out whether it can be measured independently
one option is to use gmin in models, perhaps as averages by growth form / group, but that approach will have to be tested
g0 is certainly not the same as gdark despite the fact that on average they have similar magnitudes. We know for sure gs can be much lower than fitted g0 and gdark in field conditions.
